---
title: "Introduction to JSON"
author: "Jorge Cimentada"
format: revealjs
editor: visual
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.asp = 0.618
)
```

## Introduction to JSON

-   Most APIs will return data in JavaScript Object Notation (JSON)

-   Format designed to share data over the internet

-   JSON is text-based: can be opened with any code editor

-   It supports the usual numeric/string values + arrays

-   Important: allows to build complex hierarchical structures

## Introduction to JSON

Why is JSON important? Why not use a CSV, for example? or XML? I'll skip the boring stuff and leave you with three:

1.  It's lightweight and easy to share
2.  It has support in nearly all programming languages
3.  It allows you to make nested relationships and they're easy to read

JSONs are easy to read directly from the source file. In this session we'll focus on real-world examples of the problems you'll face while working with JSONs

## Your first JSON example

```{r}
library(jsonlite)
library(tibble)
library(tidyr)
```

    {
      "president": "Parlosi",
      "vicepresident": "Kantos",
      "opposition": "Pitaso"
    }

-   Starts with `{}`

-   It's based around `key:value` pairs

-   Each `key:value` can contain numbers od strings (for now)

-   That's it. That's a JSON file.

## Your first JSON example

```{r}
json_str <- '{
  "president": "Parlosi",
  "vicepresident": "Kantos",
  "opposition": "Pitaso"
}'

fromJSON(json_str)
```

-   The `key` must be a string

-   Parsed as a named list

-   Think of JSONs as named arrays

## Real JSON example

    {
        "president": [
            {
                "last_name": "Parlosi",
                "party": "Free thinkers",
                "age": 35
            }
        ],
        "vicepresident": [
            {
                "last_name": "Kantos",
                "party": "Free thinkers",
                "age": 52
            }
        ],
        "opposition": [
            {
                "last_name": "Pitaso",
                "party": "Everyone United",
                "age": 45
            }
        ]
    }

## Real JSON example

Let's break it down:

    "president": [
        {
            "last_name": "Parlosi",
            "party": "Free thinkers",
            "age": 35
        }
    ]

-   Same JSON rules

-   Now we have an array

-   Structure: `key` followed by an array with three `key:value` pairs

## Real JSON example

-   Think of arrays in JSON as rows in a data frame.

-   Three keys (names of each slot) where each contains a 1 row data frame inside.

<!-- -->

    {
       "key":[
          ## First row
          {
             "col1":1,
             "col2":2
          },
          ## Second row
          {
             "col1":3,
             "col2":4
          }
       ]
    }

## Real JSON example

```{r, eval = FALSE}
json_str <- '
{
    "president": [
        {
            "last_name": "Parlosi",
            "party": "Free thinkers",
            "age": 35
        }
    ],
    "vicepresident": [
        {
            "last_name": "Kantos",
            "party": "Free thinkers",
            "age": 52
        }
    ],
    "opposition": [
        {
            "last_name": "Pitaso",
            "party": "Everyone United",
            "age": 45
        }
    ]
}
'

fromJSON(json_str, simplifyDataFrame = TRUE)
```

How is it parsed?

## Real JSON example

A named list of data frames:

```{r, echo = FALSE}
json_str <- '
{
    "president": [
        {
            "last_name": "Parlosi",
            "party": "Free thinkers",
            "age": 35
        }
    ],
    "vicepresident": [
        {
            "last_name": "Kantos",
            "party": "Free thinkers",
            "age": 52
        }
    ],
    "opposition": [
        {
            "last_name": "Pitaso",
            "party": "Everyone United",
            "age": 45
        }
    ]
}
'

fromJSON(json_str, simplifyDataFrame = TRUE)
```

JSON's are `key:value` pairs and the value it self can be an array with other key/value pairs.

## Real JSON example

With that explained, how will this be parsed into R?

    {
        "president": [
            {
                "last_name": "Parlosi",
                "party": "Free thinkers",
                "age": 35
            },
            {
                "last_name": "Stevensson",
                "party": "Free thinkers"
            }
        ],
        "vicepresident": [
            null
        ],
        "opposition": {
            "last_name": "Pitaso",
            "party": "Everyone United",
            "age": 45
        }
    }

## Real JSON example

Write it down before next slide.

## Real JSON example

```{r}
json_str <- '
{
    "president": [
        {
            "last_name": "Parlosi",
            "party": "Free thinkers",
            "age": 35
        },
        {
            "last_name": "Stevensson",
            "party": "Free thinkers"
        }
    ],
    "vicepresident": [
        null
    ],
    "opposition": {
        "last_name": "Pitaso",
        "party": "Everyone United",
        "age": 45
    }
}
'

res <- fromJSON(json_str)
```

## Real JSON example

```{r, echo = FALSE}
res
```

## Real JSON example

1.  A data frame for the first slot. That's correct because the JSON contained an array with two sets of `key:value` pairs. That's translatable to a data frame with two rows even though one of the two sets did not have a field for `age`.

2.  An NA value since `null` is the way missing values are represented in JSON but notice that this `null` is in a JSON array so this is effectively a data frame with an `NA` value.

3.  A named list. That's right because there is no array structure (`[...]`) so `key:value` pairs are interpreted as a named list.

## Real JSON example

Why is this important? You can fix stuff:

```{r}
res$opposition <- data.frame(res$opposition)
res
```

-   They most important tool for working with JSONs is subsetting

-   The dirtiest part of JSON is complex nesting

## The `enframe` + `unnest` strategy

Most important part of this section. Taking the previous example, how can we turn it into:

    ## # A tibble: 4 × 4
    ##   name          last_name  party             age
    ##   <chr>         <chr>      <chr>           <int>
    ## 1 president     Parlosi    Free thinkers      35
    ## 2 president     Stevensson Free thinkers      NA
    ## 3 vicepresident <NA>       <NA>               NA
    ## 4 opposition    Pitaso     Everyone United    45

-   This is the ideal summary of the result: everything in a single data frame with complete and incomplete information.

-   Each 'category' (`president`, `vicepresident`, `opposition`) is a row.

## The `enframe` + `unnest` strategy

-   General strategy of combining two functions: `enframe` and `unnest`.

-   `enframe` takes a named list and does two things:

    -   Extracts the *names* of each slot in the list and stores it in a column in a data frame.

    -   Takes **everything** inside each slot and stores it in a *list-column*.

## The `enframe` + `unnest` strategy

-   List-column: column of class list that can contain different things (if you remember, all columns in R must be of the same kind, either `numeric`, `character` or something else but there can't be two types in the same column).

-   In our example,

    -   First row of this list-column is a data frame that has two rows

    -   The second row is an empty data frame with an `NA` value

    -   The third is now a data frame since we altered the JSON manually (remember?).

## The `enframe` + `unnest` strategy

Let's take it for a spin:

```{r}
res %>%
  enframe()
```

1.  First column contains the names of the named list and the second column contains the list-column.
2.  List column has two data frame and an empty slot
3.  This structure is useful for reading nested stuff

## The `enframe` + `unnest` strategy

How do we transform it?

-   `unnest` takes list-columns and 'unpacks' them into the common class of the list.

-   If the all objects are of different classes (`data.frame`, `vectors`, etc..), then `unnest` will fail.

-   If all objects within the list are of the same class, it will combine all of them into a proper column or 'unpack' it's values.

All of this sounds convoluted right? Let's see some applied examples.

## The `enframe` + `unnest` strategy

```{r}
res %>%
  enframe() %>%
  unnest(cols = value)
```

All objects of the data frame were compatible so it combines them all.

## The `enframe` + `unnest` strategy

A failed example:

```{r}

json_str <- '
{
    "president": [
        {
            "last_name": "Parlosi",
            "party": "Free thinkers",
            "age": 35
        },
        {
            "last_name": "Stevensson",
            "party": "Free thinkers"
        }
    ],
    "vicepresident": [
        null
    ],
    "opposition": {
        "last_name": "Pitaso",
        "party": "Everyone United",
        "age": 45
    }
}
'

res <- fromJSON(json_str)
```

Can you tell me why?

## The `enframe` + `unnest` strategy

```{r, error = TRUE}
res %>%
  enframe() %>%
  unnest(cols = value)
```

<br>

The main problem you'll encounter with JSON's is that you're trying to parse some JSON that has many **nested arrays** and some of these arrays are not **compatible for unnesting** so you'll have to submerge yourself into these nested arrays and fix whatever data you want to extract.

## Accessing deeply nested JSONs

<br>

> Suppose that as part of a research project, you've recently been granted access to the API of a company. You're interested in studying the relationship between the geo-location of clients, their shopping patterns and their social class.

## Accessing deeply nested JSONs

    {
        "client1": [
            {
                "name": "Kurt Rosenwinkel",
                "device": [
                    {
                        "type": "iphone",
                        "location": [
                            {
                                "lat": 213.12,
                                "lon": 43.213
                            }
                        ]
                    }
                ]
            }
        ],
        "client2": [
            {
                "name": "YEITUEI",
                "device": [
                    {
                        "type": "android",
                        "location": [
                            {
                                "lat": 211.12,
                                "lon": 53.213
                            }
                        ]
                    }
                ]
            }
        ]
    }

## Accessing deeply nested JSONs

Before understand in-depth a JSON, it's not a terrible idea to try to read it:

```{r}
json_str <- '
{
    "client1": [
        {
            "name": "Kurt Rosenwinkel",
            "device": [
                {
                    "type": "iphone",
                    "location": [
                        {
                            "lat": 213.12,
                            "lon": 43.213
                        }
                    ]
                }
            ]
        }
    ],
    "client2": [
        {
            "name": "YEITUEI",
            "device": [
                {
                    "type": "android",
                    "location": [
                        {
                            "lat": 211.12,
                            "lon": 53.213
                        }
                    ]
                }
            ]
        }
    ]
}
'

res <- fromJSON(json_str)
```

## Accessing deeply nested JSONs

```{r, echo = FALSE}
res
```

<br>

Two data frames parsed, one for each client. Were they parsed correctly?

<br>

```{r}
res$client1
```

## Accessing deeply nested JSONs

Say we wanted to append "Device:" to the device of each client:

```{r, error = TRUE}
paste0("Device: ", res$client1$device)
```

<br> Not parsed correctly?

<br>

```{r}
res$client1$device[[1]]
```

`device` is a list with a data frame inside...

## Accessing deeply nested JSONs

```{r}
lapply(res$client1$device[[1]], class)
```

<br>

`location` is also a list. Let's look at what's inside:

```{r}
res$client1$device[[1]]$location[[1]]
```

-   General pattern: nested structured

## Accessing deeply nested JSONs

-   Whenever `fromJSON` encounters an array within an array, it converts each one to a data frame.

-   The problem is that these are recurrent data frames inside data frames and so on.

-   It's difficult to assess where to stop and this is a good example where `unnest` is very handy.

```{r}
res$client1 %>%
  as_tibble() %>%
  unnest(cols = device) %>%
  unnest(cols = location)
```

## Accessing deeply nested JSONs

Let's apply this to `res` in general:

```{r}
all_res <-
  res %>%
  enframe(name = "client")

all_res
```

<br>

```{r}
all_res %>%
  unnest(cols = "value")
```

## Accessing deeply nested JSONs

```{r}
all_res %>%
  unnest(cols = "value") %>%
  unnest(cols = "device")
```

<br>

```{r}
all_res %>%
  unnest(cols = "value") %>%
  unnest(cols = "device") %>%
  unnest(cols = "location")
```

## Summary

-   Arrays are interpreted as 'rows' in JSONs

-   Subsetting is very important for cleaning JSONs

-   Nested JSONs are your biggest problem

-   `enframe` + `unnest` strategy can help you with nested JSONs

-   Problems: objects in a `list-column` are of different classes

## Homework

-   Chapter 14 -- read + exercises

-   Everyone should have a project + a group [here](https://docs.google.com/spreadsheets/d/11QmdgRXvbwtN9hSOwVyw9-iVjrN-yeRq_BOudidAnX8/edit#gid=73208201).

-   Work on project should begin now. Submissions can be done from now on. Final deadline for the project is in two weeks.
